---
title: "PA1_template.Rmd"
author: "Duncan Simon"
date: "9 November 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(lattice)
```

## Loading and preprocessing the data

```{r}
activityData <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
# Total steps per day as histogram; mean and median reported
stepsPerDay <- aggregate(steps ~ date, activityData, sum)
hist(stepsPerDay$steps, main="Total Steps Per Day", xlab="Number of Steps")
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
```
* Mean: `r meanSteps`
* Median: `r medianSteps`

## What is the average daily activity pattern?
```{r}
# Average daily pattern activity; max steps reported
meanActivity <- aggregate(steps ~ interval, activityData, mean)
plot(meanActivity$interval, meanActivity$steps, type="l", main="Average number of steps per time interval", ylab="Mean Steps", xlab="Interval (Minutes)")
maxInterval <- meanActivity[which.max(meanActivity$steps), 1]
```
* Most steps at interval: `r maxInterval`

## Imputing missing values
```{r}
# Number of NA rows
NARows <- sum(is.na(activityData$steps))
```
* Number of NA rows: `r NARows`

```{r}
# Fill NAs with mean for that interval, plot total steps per day as histogram
# Mean and median reported
imputedActivityData <- transform(activityData, steps=ifelse(is.na(activityData$steps), meanActivity$steps[match(activityData$interval, meanActivity$interval)], activityData$steps))
imputedStepsPerDay <- aggregate(steps ~ date, imputedActivityData, sum)
hist(imputedStepsPerDay$steps, main="Total Steps Per Day", xlab="Number of Steps")
meanStepsImputed <- mean(imputedStepsPerDay$steps)
medianStepsImputed <- median(imputedStepsPerDay$steps)
```
* Mean (imputed): `r meanStepsImputed`
* Median (imputed): `r medianStepsImputed`

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## Differences in weekday and weekend activity patterns
weekendDays <- c("Saturday", "Sunday")
imputedActivityData$weekday <- as.factor(ifelse(is.element(weekdays(as.Date(imputedActivityData$date)), weekendDays), "Weekend", "Weekday"))
imputedMeanActivity <- aggregate(steps ~ interval + weekday, imputedActivityData, mean)
xyplot(imputedMeanActivity$steps ~ imputedMeanActivity$interval|imputedMeanActivity$weekday, main="Average Steps per Day Type, by Interval",xlab="Interval", ylab="Number of Steps",layout=c(1,2), type="l")
```